---
title: "Braindata"
author: "Zhonghan Pei"
date: "4/10/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Read the data
```{r,warning=FALSE,echo=FALSE}
library(readr)
brain_data <- read_csv("RSPW.csv")
library(gdata)
clinc<-read.xls("20190220_Clinc.xlsx",sheet = 2,header = T)
```

```{r,echo=FALSE}
library(dplyr)
new_clinc <- clinc[,c(1,5,6,7,33,34,35,36,37,38,39,40,41,42,43,47,48)]
new_clinc[31,10] = 0
new_clinc$Veg <- rowSums(new_clinc[,c(9,10,11,12,13,14,15)])
new_clinc$Amc <- rowSums(new_clinc[,c(5,6)])
new_clinc$Med <- rowSums(new_clinc[,c(7,8)])

new_clinc$diet <- ifelse(new_clinc$Veg == 1 | new_clinc$Veg == 2, "Veg", ifelse((new_clinc$Amc == 1 | new_clinc$Amc == 2) & (new_clinc$Amc > new_clinc$Med), "Amc", ifelse(new_clinc$Med == 1 | new_clinc$Med == 2, "Med", NA)))

new_clinc$diet[180] = "Amc"
new_clinc$diet[164] = "Med"
new_clinc$diet[155] = "Med"
new_clinc$diet[142] = "Med"
new_clinc$diet[126] = "Amc"
new_clinc$diet[117] = "Med"
new_clinc$diet[90] = "Amc"
new_clinc$diet[67] = "Med"
new_clinc$diet[65] = "Med"
new_clinc$diet[56] = "Amc"
new_clinc$diet[18] = "Med"

new_clinc$obs <- ifelse(new_clinc$BMI < 25, "Lean", "Over")
#ifelse((new_clinc$BMI >= 25 & new_clinc$BMI < 30) , "Over", "Obese"))

new_clinc <- new_clinc[order(new_clinc$NDPNum),]
col_names <- colnames(new_clinc)
col_names1 <- colnames(brain_data)[-1]
new_clinc1 <- merge(new_clinc,brain_data, by.x="NDPNum", by.y="Subject_ID")
new_clinc <- new_clinc1[,col_names]
new_brain <- new_clinc1[,col_names1]
```

Brainstem
Nucleus accumbens (NACC)
Putamen (Put)
Caudate Nucleus (CaN)
Globus Pallidum (Pal)
```{r,echo=FALSE}
library(dplyr)
library(stringr)
strings <- c("B_Bstem","Pal","CaN","Pu","Nacc")
new_brain <- new_brain[,str_detect(colnames(new_brain), paste(strings, collapse = "|"))]

set.seed(12345)
training_sample <- sample(nrow(new_brain),nrow(new_brain)*0.7)
new_brain_train <- new_brain[training_sample,]
new_brain_test <- new_brain[-training_sample,]
```

```{r,echo=FALSE}
library(ggbiplot)
##Brain PCA 
pca_brain <- prcomp(new_brain,scale = TRUE)
ggbiplot(pca_brain,choices=c(1,2),obs.scale = 1, var.scale = 1,groups = new_clinc$obs,ellipse = TRUE,var.axes=FALSE)

pca_brain <- prcomp(new_brain_train,scale = TRUE)
summ_pca<-summary(pca_brain)
```

##Loadings and Principal components
```{r}
##Percentage of variances explained by our 20 principal components
sum(summ_pca$importance[2,1:20])

##Get the rotation matrix or loadings from pca results
rotation_matrix <- pca_brain$rotation[,1:20]

##Convert to matrix form
matrix_train <- matrix(as.numeric(unlist(new_brain_train)),nrow =129)

##New matrix with 20 principal components
train_PC20 <- matrix_train%*%rotation_matrix
train_PC20 <- as.data.frame(train_PC20)
diet_train <- new_clinc$diet[training_sample]

##Predict diet using the 20 principal components we collect

library(nnet)
glm.fit = multinom(diet_train~.,data = train_PC20)
summary(glm.fit)

##Loadings for PC1
head(rotation_matrix[order(rotation_matrix[,1]),1])

tail(rotation_matrix[order(rotation_matrix[,1]),1])

##Loadings for PC2
head(rotation_matrix[order(rotation_matrix[,2]),2])

tail(rotation_matrix[order(rotation_matrix[,2]),2])
```

##Lets predict our testing data sets using logistic regression model we have 
```{r,echo=FALSE}
##Get the testing data
matrix_test <- matrix(as.numeric(unlist(new_brain_test)),nrow =nrow(new_brain_test))

test_PC20 <- matrix_test%*%rotation_matrix
test_PC20 <- as.data.frame(test_PC20)


##Predicted values for testing set
predicted_test <- glm.fit %>% predict(test_PC20)
diet_test<- new_clinc$diet[-training_sample]

##Cross Validation score
sum(predicted_test == new_clinc$diet[-training_sample])/length(predicted_test)


##Predicted values for training set
predicted_train <- glm.fit %>% predict(train_PC20)

sum(predicted_train == diet_train)/length(predicted_train)
```

##Lets predict our testing data using KNN
```{r,echo=FALSE}
knn_train_PC20 <- cbind(train_PC20,diet = diet_train)
knn_test_PC20 <- cbind(test_PC20, diet = diet_test)
```

```{r}
library(DMwR)
knn_test <-kNN(diet~.,knn_train_PC20,knn_test_PC20, norm = TRUE, k =5)
sum(knn_test==diet_test)/length(knn_test)
```

##Lets predict our result in random forest
```{r}
library(randomForest)
set.seed(12345)
rf <- randomForest(diet~., data = knn_train_PC20, ntree = 500, mtry = 10)
##Prediction model test for training set
rf_predicted <- predict(rf, knn_train_PC20,type = "class")
table(rf_predicted, diet_train)

##Prediction model test for testing set
rf_predicted_test <- predict(rf, knn_test_PC20,type = "class")
table(rf_predicted_test,diet_test)

sum(rf_predicted_test == diet_test)/length(diet_test)
```

##Lets predict our result in neural network
```{r}
library(nnet)
set.seed(12345)
nn <- nnet(diet~., data = knn_train_PC20, maxit = 150,size = 4)
##Prediction model test for training set
table(predict(nn, knn_train_PC20,type = "class"),diet_train)
sum(predict(nn, knn_train_PC20,type = "class") == diet_train)/(length(diet_train))

##Prediction model test for testing set
table(predict(nn, knn_test_PC20,type = "class"),diet_test)
sum(predict(nn, knn_test_PC20,type = "class") == diet_test)/(length(diet_test))

```

##Prediction Using 40 PCs
```{r}
##Percentage of variances explained by our 40 principal components
sum(summ_pca$importance[2,1:40])

##Get the rotation matrix or loadings from pca results
rotation_matrix_40 <- pca_brain$rotation[,1:40]

##Convert to matrix form
matrix_train <- matrix(as.numeric(unlist(new_brain_train)),nrow =129)

##New matrix with 20 principal components
train_PC40 <- matrix_train%*%rotation_matrix_40
train_PC40 <- as.data.frame(train_PC40)
diet_train_40 <- new_clinc$diet[training_sample]

##Predict diet using the 20 principal components we collect

library(nnet)
glm.fit = multinom(diet_train_40~.,data = train_PC40)
summary(glm.fit)
```

##Lets predict our testing data sets using logistic regression model we have 
```{r}
##Get the testing data
matrix_test <- matrix(as.numeric(unlist(new_brain_test)),nrow =nrow(new_brain_test))

test_PC40 <- matrix_test%*%rotation_matrix_40
test_PC40 <- as.data.frame(test_PC40)


##Predicted values for testing set
predicted_test_40 <- glm.fit %>% predict(test_PC40)
diet_test_40<- new_clinc$diet[-training_sample]


##Predicted values for training set
predicted_train_40 <- glm.fit %>% predict(train_PC40)

sum(predicted_train == diet_train)/length(predicted_train_40)

##Cross Validation score
sum(predicted_test_40 == new_clinc$diet[-training_sample])/length(predicted_test_40)
```

##Lets predict our testing data using KNN
```{r}
knn_train_PC40 <- cbind(train_PC40,diet = diet_train_40)
knn_test_PC40 <- cbind(test_PC40, diet = diet_test_40)

library(DMwR)
knn_test <-kNN(diet~.,knn_train_PC40,knn_test_PC40, norm = TRUE, k =5)
sum(knn_test==diet_test_40)/length(knn_test)
```

##Lets predict our result in random forest
```{r}
library(randomForest)
set.seed(12345)
rf <- randomForest(diet~., data = knn_train_PC40, ntree = 600, mtry = 10)
##Prediction model test for training set
rf_predicted <- predict(rf, knn_train_PC40,type = "class")
table(rf_predicted, diet_train_40)

##Prediction model test for testing set
rf_predicted_test <- predict(rf, knn_test_PC40,type = "class")
table(rf_predicted_test,diet_test)

sum(rf_predicted_test == diet_test_40)/length(diet_test_40)

```

```{r}
library(nnet)
set.seed(12345)
nn <- nnet(diet~., data = knn_train_PC40, maxit = 150,size = 4)
##Prediction model test for training set
table(predict(nn, knn_train_PC40,type = "class"),diet_train_40)
sum(predict(nn, knn_train_PC40,type = "class") == diet_train_40)/(length(diet_train_40))

##Prediction model test for testing set
table(predict(nn, knn_test_PC40,type = "class"),diet_test_40)
sum(predict(nn, knn_test_PC40,type = "class") == diet_test_40)/(length(diet_test_40))
```

##Brain PCA 
```{r}
pca_brain <- prcomp(new_brain,scale = TRUE)
ggbiplot(pca_brain,choices=c(1,2),obs.scale = 1, var.scale = 1,groups = new_clinc$obs,ellipse = TRUE,var.axes=FALSE)
```

Hiereachical Analysis
```{r,echo=FALSE}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering visualization
library(dendextend) # for comparing two dendrograms
scale_new_brain <- scale(new_brain)

# Dissimilarity matrix
d1 <- dist(scale_new_brain, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d1, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)

##Amc Group
amc_brain <- new_brain[which(new_clinc$diet=="Amc"),]
scale_amc_brain <- scale(amc_brain)

# Dissimilarity matrix
d2 <- dist(scale_amc_brain, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d2, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)

##Med Group
med_brain <- new_brain[which(new_clinc$diet=="Med"),]
scale_med_brain <- scale(med_brain)

# Dissimilarity matrix
d3 <- dist(scale_med_brain, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d3, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)

##Veg Group
veg_brain <- new_brain[which(new_clinc$diet=="Veg"),]
scale_veg_brain <- scale(veg_brain)

# Dissimilarity matrix
d4 <- dist(scale_veg_brain, method = "euclidean")

# Hierarchical clustering using Complete Linkage
hc1 <- hclust(d4, method = "complete" )

# Plot the obtained dendrogram
plot(hc1, cex = 0.6, hang = -1)
```

##Lets deal with our food addiction table
```{r}
library(stringr)
YFAS_new <- clinc[,c("NDPNum","Sex","BMI","YFAS_SymptomCount")]

YFAS_new <- merge(YFAS_new, new_clinc[,c("NDPNum","diet","obs")],by = "NDPNum")

YFAS_new1 <- merge(YFAS_new,brain_data, by.x="NDPNum", by.y="Subject_ID")
YFAS_new1 <- YFAS_new1[complete.cases(YFAS_new1),]
YFAS <- YFAS_new1[,c("Sex","BMI","YFAS_SymptomCount","obs","diet")]
YFAS$YFhigh <- ifelse(YFAS$YFAS_SymptomCount < 2, "low","high")

new_brain_Y <- YFAS_new1[,str_detect(colnames(YFAS_new1), paste(strings, collapse = "|"))]
set.seed(12345)
training_sample <- sample(nrow(new_brain_Y),nrow(new_brain_Y)*0.7)
new_brain_Y_train <- new_brain_Y[training_sample,]
new_brain_Y_test <- new_brain_Y[-training_sample,]
# YFAS[complete.cases(YFAS),]
```

```{r}
library(ggbiplot)
##Brain PCA 
pca_brain <- prcomp(new_brain_Y,scale = TRUE)
ggbiplot(pca_brain,choices=c(1,2),obs.scale = 1, var.scale = 1,groups = YFAS$YFhigh,ellipse = TRUE,var.axes=FALSE)

pca_brain <- prcomp(new_brain_Y_train,scale = TRUE)
summ_pca<-summary(pca_brain)
```

```{r}
##Percentage of variances explained by our 20 principal components
sum(summ_pca$importance[2,1:20])

##Get the rotation matrix or loadings from pca results
rotation_matrix <- pca_brain$rotation[,1:20]

##Convert to matrix form
matrix_train <- matrix(as.numeric(unlist(new_brain_Y_train)),nrow =118)

##New matrix with 20 principal components
train_PC20 <- matrix_train%*%rotation_matrix
train_PC20 <- as.data.frame(train_PC20)
YFAS_train <- YFAS_new$YFhigh[training_sample]

##Predict diet using the 20 principal components we collect

library(nnet)
glm.fit = multinom(YFAS_train~.,data = train_PC20)
summary(glm.fit)
```

```{r}
##Get the testing data
matrix_test <- matrix(as.numeric(unlist(new_brain_Y_test)),nrow =nrow(new_brain_Y_test))

test_PC20 <- matrix_test%*%rotation_matrix
test_PC20 <- as.data.frame(test_PC20)


##Predicted values for testing set
predicted_test <- glm.fit %>% predict(test_PC20)
YFAS_test <- YFAS_new$YFhigh[-training_sample]

##Cross Validation score
sum(predicted_test == YFAS_test)/length(predicted_test)


##Predicted values for training set
predicted_train <- glm.fit %>% predict(train_PC20)

sum(predicted_train == YFAS_train)/length(predicted_train)
```

##Lets use KNN
```{r}
knn_train_PC20 <- cbind(train_PC20,YFAS = YFAS_train)
knn_test_PC20 <- cbind(test_PC20, YFAS = YFAS_test)

library(DMwR)
knn_test <-kNN(YFAS~.,knn_train_PC20,knn_test_PC20, norm = TRUE, k =5)
sum(knn_test==YFAS_test)/length(knn_test)
```

##Lets use random forest
```{r,}
library(randomForest)
set.seed(12345)
rf <- randomForest(YFAS~., data = knn_train_PC20, ntree = 500, mtry = 10)
##Prediction model test for training set
rf_predicted <- predict(rf, knn_train_PC20,type = "class")
table(rf_predicted, YFAS_train)

##Prediction model test for testing set
rf_predicted_test <- predict(rf, knn_test_PC20,type = "class")
table(rf_predicted_test,YFAS_test)

sum(rf_predicted_test == YFAS_test)/length(YFAS_test)
```