---
title: "Exploratory Analysis"
author: "Zhonghan Pei"
date: "2/27/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Read the Data Set
```{r, echo=FLASE}
library(gdata)
clinc<-read.xls("20190220_Clinc.xlsx",sheet = 2,header = T)
library(readr)
brain_data <- read_csv("RSPW.csv")
```

##Extract Age, Sex, BMI, Diet Data
```{r,echo=FALSE}
library(dplyr)
new_clinc <- clinc[,c(5,6,7,33,34,35,36,37,38,39,40,41,42,43,47,48)]
new_clinc[31,9] = 0
new_clinc$Veg <- rowSums(new_clinc[,c(8,9,10,11,12,13,14)])
new_clinc$Amc <- rowSums(new_clinc[,c(4,5)])
new_clinc$Med <- rowSums(new_clinc[,c(6,7)])

new_clinc$diet <- ifelse(new_clinc$Veg == 1 | new_clinc$Veg == 2, "Veg", ifelse((new_clinc$Amc == 1 | new_clinc$Amc == 2) & (new_clinc$Amc > new_clinc$Med), "Amc", ifelse(new_clinc$Med == 1 | new_clinc$Med == 2, "Med", NA)))

new_clinc$diet[180] = "Amc"
new_clinc$diet[164] = "Med"
new_clinc$diet[155] = "Med"
new_clinc$diet[142] = "Med"
new_clinc$diet[126] = "Amc"
new_clinc$diet[117] = "Med"
new_clinc$diet[90] = "Amc"
new_clinc$diet[67] = "Med"
new_clinc$diet[65] = "Med"
new_clinc$diet[56] = "Amc"
new_clinc$diet[18] = "Med"

new_clinc$obs <- ifelse(new_clinc$BMI < 25, "Lean", ifelse((new_clinc$BMI >= 25 & new_clinc$BMI < 30) , "Over", "Obese"))

#count(new_clinc[new_clinc$Sex == 1,])
#count(new_clinc[new_clinc$Sex == 2,])

library(dplyr)
set.seed(12345)
new_sample <- rbind(new_clinc[new_clinc$Sex == 1,],sample_n(new_clinc[new_clinc$Sex == 2,],66))
```
Binomial P-value Test
```{r}
##Sexuality by ratio
#sum(47+17+2)/185
##Test if American standard diet has sexual perferences
#pbinom(47, size=(47+85), prob=(sum(47+17+2)/185))
##Test if Mediterrean diet has sexual perferences
#pbinom(17, size = 34, prob = (sum(47+17+2)/185))
##Test if Vegetarian diet has sexual perferences
#pbinom(2, size = 19, prob = (sum(47+17+2)/185))

bi_test1 <- binom.test(47,(47+85),(sum(47+17+2)/185),alternative="two.sided")
bi_test1$p.value

bi_test2 <- binom.test(17,(34),(sum(47+17+2)/185),alternative="two.sided")
bi_test2$p.value

bi_test3 <- binom.test(2,(19),(sum(47+17+2)/185),alternative="two.sided")
bi_test3$p.value
```

Table of all groups
```{r}
num1 <- c(nrow(new_clinc[(new_clinc$Sex == 1 & new_clinc$diet == "Amc"),]),nrow(new_clinc[(new_clinc$Sex == 2 & new_clinc$diet == "Amc"),]))

num2 <- c(nrow(new_clinc[(new_clinc$Sex == 1 & new_clinc$diet == "Med"),]),nrow(new_clinc[(new_clinc$Sex == 2 & new_clinc$diet == "Med"),]))

num3 <- c(nrow(new_clinc[(new_clinc$Sex == 1 & new_clinc$diet == "Veg"),]),nrow(new_clinc[(new_clinc$Sex == 2 & new_clinc$diet == "Veg"),]))

diet_1 <- c(rep(c("Amc"),2),rep(c("Med"),2),rep(c("Veg"),2))

sex_1 <- c(rep(c("Male","Female"),3))

num <- c(num1, num2, num3)

data1 <- data.frame(num,diet_1,sex_1)
data1


```

From the results above we can reject the null hypothesis that the sexuality has no difference on dietary choices.Therefore, sexuality does affect our dietary choices especially in Vegetarian diet.

##Boostrapping Test on Obesity and Diet

Table of all groups
```{r}
num1 <- c(nrow(new_clinc[(new_clinc$obs == "Obese" & new_clinc$diet == "Amc"),]),nrow(new_clinc[(new_clinc$obs == "Over" & new_clinc$diet == "Amc"),]),nrow(new_clinc[(new_clinc$obs == "Lean" & new_clinc$diet == "Amc"),]))

num2 <- c(nrow(new_clinc[(new_clinc$obs == "Obese" & new_clinc$diet == "Med"),]),nrow(new_clinc[(new_clinc$obs == "Over" & new_clinc$diet == "Med"),]),nrow(new_clinc[(new_clinc$obs == "Lean" & new_clinc$diet == "Med"),]))

num3 <-c(nrow(new_clinc[(new_clinc$obs == "Obese" & new_clinc$diet == "Veg"),]),nrow(new_clinc[(new_clinc$obs == "Over" & new_clinc$diet == "Veg"),]),nrow(new_clinc[(new_clinc$obs == "Lean" & new_clinc$diet == "Veg"),]))

diet_1 <- c(rep(c("Amc"),3),rep(c("Med"),3),rep(c("Veg"),3))

obs_1 <- c(rep(c("Obese","Over","Lean"),3))

num <- c(num1, num2, num3)

data <- data.frame(num,diet_1,obs_1)
data


```

Collect BMI datasets 

```{r}
BMI_data <-data.frame(BMI = new_clinc$BMI,diet = new_clinc$diet)

bootstrapping <- function(x,y,R){
  ##Calculate original test statistic
  t_orig = t.test(x,y, alternative = "two.sided")$statistic
  x_1 <- x - mean(x) + mean(c(x,y))
  y_1 <- y - mean(y) + mean(c(x,y))
  ##Create Null vector
  test_stats <- NULL
  ##Test statistic for each iteration
  for (i in 1:R) {
    x_s <- sample(x_1, size = length(x),replace = TRUE)
    y_s <- sample(y_1,size = length(y),replace = TRUE)
    test_stats<-c(test_stats,t.test(x_s,y_s, alternative = "two.sided")$statistic)
  }
  ##Return the p-value for testing equal mean
  return(sum(test_stats > t_orig)/length(test_stats))
}

BMI_amc <- BMI_data[BMI_data$diet == "Amc",1]

BMI_med <- BMI_data[BMI_data$diet == "Med",1]

BMI_veg <- BMI_data[BMI_data$diet == "Veg",1]


summary(BMI_amc)
summary(BMI_med)
summary(BMI_veg)
bootstrapping(BMI_amc,BMI_med, 1000)
t.test(BMI_amc,BMI_med, alternative = "two.sided")$statistic
bootstrapping(BMI_amc,BMI_veg, 1000)
t.test(BMI_amc,BMI_veg, alternative = "two.sided")$statistic
bootstrapping(BMI_med,BMI_veg, 1000)
t.test(BMI_med,BMI_veg, alternative = "two.sided")$statistic


hist(BMI_data$BMI)
```

From the results that we can conclude that in our data patients taking American standard diet have significantly different BMI than those who take Mediterrean diet.

```{r}
par(mfrow = c(3,1))
hist(BMI_amc,col = "red",main = "Histogram BMI for American Standard Diet")

hist(BMI_med,col="blue",main = "Histogram BMI for Mediterrean Diet")

hist(BMI_veg,col = "green",main = "Histogram BMI for Vegetarian Diet")
```

##Food addiction 
```{r}
YFAS <- clinc[,c("Sex","BMI","YFAS_SymptomCount")]
YFAS$YFhigh <- ifelse(YFAS$YFAS_SymptomCount < 2, "low","high")
YFAS <- YFAS[complete.cases(YFAS),]
# nrow(YFAS[YFAS$Sex == 1,]) 60
# nrow(YFAS[YFAS$Sex == 2,]) 110
# nrow(YFAS[YFAS$YFhigh == "high",]) 58
# nrow(YFAS[YFAS$YFhigh == "low",]) 112
# nrow(YFAS[YFAS$YFhigh == "high" & YFAS$Sex == 1,]) 16
# nrow(YFAS[YFAS$YFhigh == "high" & YFAS$Sex == 2,]) 42
# nrow(YFAS[YFAS$YFhigh == "low" & YFAS$Sex == 1,]) 44
# nrow(YFAS[YFAS$YFhigh == "low" & YFAS$Sex == 2,]) 68
bi_test1 <- binom.test(16,(16+42),(60/170),alternative="two.sided")
bi_test1$p.value

bi_test2 <- binom.test(44,(44+68),(60/185),alternative="two.sided")
bi_test2$p.value

BMI_high <- YFAS$BMI[YFAS$YFhigh=="high"]

BMI_low <- YFAS$BMI[YFAS$YFhigh=="low"]

bootstrapping(BMI_high,BMI_low, 1000)
t.test(BMI_high,BMI_low, alternative = "two.sided")$statistic
```
From the above result we can clearly say that YFAS food addiction score can greatly affect BMI.


